---
title: "The Big Five Personality Test"
author: "Mine Cetinkaya-Rundel"
date: "5/29/2018"
output: html_document
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 1, digits = 2)
```

The Big Five personality traits is a theory of five broad dimensions used by some psychologists to describe the human personality and psyche: openness to experience, conscientiousness, extraversion, agreeableness, and neuroticism.[^1]

```{r message=FALSE}
big5 <- read_delim("raw-data/big5.txt", delim = "\t")
```

## Data cleaning

A data cleanup script is in the `\code` folder. Let's run that.

```{r message=FALSE, warning=FALSE}
source("code/01-data-cleanup.R")
```

## Ages

```{r message=FALSE, fig.height=3}
ggplot(big5, aes(x = age)) +
  geom_histogram()
```

```{r}
summary(big5$age)
```

## Regress extraversion vs. neuroticism and gender {.smaller}

**Extraversion:** Seeking fulfillment from sources outside the self or in community. High scorers are social, low scorers prefer to work alone. **Neuroticism:** Being emotional.

```{r message=FALSE}
m_ext_age <- lm(extraversion ~ neuroticism * gender, data = big5)
tidy(m_ext_age)
```

## Grab coefficients

```{r}
tidy_m <- tidy(m_ext_age)
```

Slope for neuroticism is `r tidy_m$estimate[2]`.


## Plot extraversion vs. age and gender

```{r message=FALSE, fig.height=4}
ggplot(data = big5, aes(x = neuroticism, y = extraversion, color = gender)) +
  geom_point(alpha = 0.5) +
  geom_jitter() +
  geom_smooth(method = "lm")
```

## Suppose you want only teens

```{r}
big5_teen <- filter(big5, age <= 19)
```

```{r}
m_ext_age_teen <- lm(extraversion ~ age * gender, data = big5_teen)
tidy(m_ext_age_teen)
```

## Plot for only teens

```{r message=FALSE, fig.height=4}
ggplot(data = big5_teen, aes(x = neuroticism, y = extraversion, color = gender)) +
  geom_point(alpha = 0.5) +
  geom_jitter() +
  geom_smooth(method = "lm")
```



[^1]: Sources: [Wikipedia](https://en.wikipedia.org/wiki/Big_Five_personality_traits) and http://personality-testing.info/_rawdata/.

